### 2022-03-16 outputs/paper/paper.Rmd
rmd, data prep

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read in the data
library(tidyverse)
table_8 <- readxl::read_excel("~/Australia-2020/inputs/data/GSS_Table8.xlsx", sheet = "Table 8.1_Mean Score")
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read in the data
library(tidyverse)
table_8_RSE <- readxl::read_excel("~/Australia-2020/inputs/data/GSS_Table8.xlsx", sheet = "Table 8.2_RSE, mean")
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

Talk way more about it.

# Results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by age
life_satis_age <- as.data.frame(table_8[8:13,])
life_satis_age <- life_satis_age[2:6,] %>% clean_names() %>%
  rename(age_group = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure1, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Age Group", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by age groups.
life_satis_age %>%
  ggplot(aes(x = age_group, y = mean_score, fill = age_group)) +
  geom_bar(position="dodge", stat="identity") +
  labs(title = "Life Satisfaction By Age",
       x = "Age Group",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by mental health condition
life_satis_mental_health <- as.data.frame(table_8[18:19,])
life_satis_mental_health <- life_satis_mental_health %>% clean_names() %>%
  rename(menal_health_condition = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure2, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Mental Health Condition", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by mental health conditions.
life_satis_mental_health %>%
  ggplot(aes(x = menal_health_condition, y = mean_score, colour = menal_health_condition, fill = menal_health_condition)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Mental Health Conditions",
       x = "Mental Health Condition",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by mental health condition
life_satis_health <- as.data.frame(table_8[21:22,])
life_satis_health <- life_satis_health %>% clean_names() %>%
  rename(health_condition = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure3, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Health Condition", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by health conditions.
life_satis_health %>%
  ggplot(aes(x = health_condition, y = mean_score, fill = health_condition)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Health Conditions",
       x = "Health Condition",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by disability
life_satis_disability <- as.data.frame(table_8[24:25,])
life_satis_disability <- life_satis_disability %>% clean_names() %>%
  rename(disability_or_not = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure4, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Disability Condition", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by disability conditions.
life_satis_disability %>%
  ggplot(aes(x = disability_or_not, y = mean_score, fill = disability_or_not)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("Yes", "No")) +
  labs(title = "Life Satisfaction By Disability Conditions",
       x = "Disability Condition",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by immigration status
life_satis_immigrate <- as.data.frame(table_8[15:16,])
life_satis_immigrate <- life_satis_immigrate %>% clean_names() %>%
  rename(recent_migrants_or_not = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure5, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Immigration Status", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by immigration status
life_satis_immigrate %>%
  ggplot(aes(x = recent_migrants_or_not, y = mean_score, fill = recent_migrants_or_not)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Immigration Status",
       x = "Recent Migrants",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract RSE of life satisfaction mean scores by immigration status
life_satis_immigrate_RSE <- as.data.frame(table_8_RSE[15:16,])
life_satis_immigrate_RSE <- life_satis_immigrate_RSE %>% clean_names() %>%
  rename(recent_migrants_or_not = australian_bureau_of_statistics, RSE_mean_score = x12)
```

```{r, Figure6, fig.cap="RSE of Mean Life Satisfaction Score of Australians in 2020 By Immigration Status", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by immigration status
life_satis_immigrate_RSE %>%
  ggplot(aes(x = recent_migrants_or_not, y = RSE_mean_score, fill = recent_migrants_or_not)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "RSE of Mean Life Satisfaction Score By Immigration Status",
       x = "Recent Migrants",
       y = "RSE of Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by sexual orientation
life_satis_sex_orien <- as.data.frame(table_8[27:28,])
life_satis_sex_orien <- life_satis_sex_orien %>% clean_names() %>%
  rename(sexual_orientation = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure7, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Sexual Orientation", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by sexual orientation.
life_satis_sex_orien %>%
  ggplot(aes(x = sexual_orientation, y = mean_score, fill = sexual_orientation)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Sexual Orientation",
       x = "Heterosexual or Not",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract RSE of life satisfaction scores by sexual orientation
life_satis_sex_orien_RSE <- as.data.frame(table_8_RSE[27:28,])
life_satis_sex_orien_RSE <- life_satis_sex_orien_RSE %>% clean_names() %>%
  rename(sexual_orientation = australian_bureau_of_statistics, RSE_of_mean = x12)
```

```{r, Figure8, fig.cap="RSE of Mean Life Satisfaction Score of Australians in 2020 By Sexual Orientation", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life RSE of satisfaction mean score by sexual orientation.
life_satis_sex_orien_RSE %>%
  ggplot(aes(x = sexual_orientation, y = RSE_of_mean, fill = sexual_orientation)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "RSE of Mean Life Satisfaction Score By Sexual Orientation",
       x = "Heterosexual or Not",
       y = "RSE of Mean Life Satisfaction Score") +
  theme_minimal()
```

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References