### 2022-03-17 outputs/paper/paper.Rmd
discuss edited

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read in the data
library(tidyverse)
table_8 <- readxl::read_excel("~/Australia-2020/inputs/data/GSS_Table8.xlsx", sheet = "Table 8.1_Mean Score")
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read in the data
library(tidyverse)
table_8_RSE <- readxl::read_excel("~/Australia-2020/inputs/data/GSS_Table8.xlsx", sheet = "Table 8.2_RSE, mean")
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

Talk way more about it.

# Results

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by age
life_satis_age <- as.data.frame(table_8[8:13,])
life_satis_age <- life_satis_age[2:6,] %>% clean_names() %>%
  rename(age_group = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure1, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Age Group", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by age groups.
life_satis_age %>%
  ggplot(aes(x = age_group, y = mean_score, fill = age_group)) +
  geom_bar(position="dodge", stat="identity") +
  labs(title = "Life Satisfaction By Age",
       x = "Age Group",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by mental health condition
life_satis_mental_health <- as.data.frame(table_8[18:19,])
life_satis_mental_health <- life_satis_mental_health %>% clean_names() %>%
  rename(menal_health_condition = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure2, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Mental Health Condition", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by mental health conditions.
life_satis_mental_health %>%
  ggplot(aes(x = menal_health_condition, y = mean_score, colour = menal_health_condition, fill = menal_health_condition)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Mental Health Conditions",
       x = "Mental Health Condition",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by mental health condition
life_satis_health <- as.data.frame(table_8[21:22,])
life_satis_health <- life_satis_health %>% clean_names() %>%
  rename(health_condition = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure3, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Health Condition", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by health conditions.
life_satis_health %>%
  ggplot(aes(x = health_condition, y = mean_score, fill = health_condition)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Health Conditions",
       x = "Health Condition",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by disability
life_satis_disability <- as.data.frame(table_8[24:25,])
life_satis_disability <- life_satis_disability %>% clean_names() %>%
  rename(disability_or_not = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure4, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Disability Condition", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by disability conditions.
life_satis_disability %>%
  ggplot(aes(x = disability_or_not, y = mean_score, fill = disability_or_not)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("Yes", "No")) +
  labs(title = "Life Satisfaction By Disability Conditions",
       x = "Disability Condition",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by immigration status
life_satis_immigrate <- as.data.frame(table_8[15:16,])
life_satis_immigrate <- life_satis_immigrate %>% clean_names() %>%
  rename(recent_migrants_or_not = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure5, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Immigration Status", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by immigration status
life_satis_immigrate %>%
  ggplot(aes(x = recent_migrants_or_not, y = mean_score, fill = recent_migrants_or_not)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Immigration Status",
       x = "Recent Migrants",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract RSE of life satisfaction mean scores by immigration status
life_satis_immigrate_RSE <- as.data.frame(table_8_RSE[15:16,])
life_satis_immigrate_RSE <- life_satis_immigrate_RSE %>% clean_names() %>%
  rename(recent_migrants_or_not = australian_bureau_of_statistics, RSE_mean_score = x12)
```

```{r, Figure6, fig.cap="RSE of Mean Life Satisfaction Score of Australians in 2020 By Immigration Status", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by immigration status
life_satis_immigrate_RSE %>%
  ggplot(aes(x = recent_migrants_or_not, y = RSE_mean_score, fill = recent_migrants_or_not)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "RSE of Mean Life Satisfaction Score By Immigration Status",
       x = "Recent Migrants",
       y = "RSE of Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract life satisfaction mean scores by sexual orientation
life_satis_sex_orien <- as.data.frame(table_8[27:28,])
life_satis_sex_orien <- life_satis_sex_orien %>% clean_names() %>%
  rename(sexual_orientation = australian_bureau_of_statistics, mean_score = x12)
```

```{r, Figure7, fig.cap="Mean Life Satisfaction Score of Australians in 2020 By Sexual Orientation", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life satisfaction mean score by sexual orientation.
life_satis_sex_orien %>%
  ggplot(aes(x = sexual_orientation, y = mean_score, fill = sexual_orientation)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "Life Satisfaction By Sexual Orientation",
       x = "Heterosexual or Not",
       y = "Mean Life Satisfaction Score") +
  theme_minimal()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Extract RSE of life satisfaction scores by sexual orientation
life_satis_sex_orien_RSE <- as.data.frame(table_8_RSE[27:28,])
life_satis_sex_orien_RSE <- life_satis_sex_orien_RSE %>% clean_names() %>%
  rename(sexual_orientation = australian_bureau_of_statistics, RSE_of_mean = x12)
```

```{r, Figure8, fig.cap="RSE of Mean Life Satisfaction Score of Australians in 2020 By Sexual Orientation", echo=FALSE, warning=FALSE, message=FALSE}
# Draw bar plot showing life RSE of satisfaction mean score by sexual orientation.
life_satis_sex_orien_RSE %>%
  ggplot(aes(x = sexual_orientation, y = RSE_of_mean, fill = sexual_orientation)) +
  geom_bar(position="dodge", stat="identity") +
  scale_x_discrete(labels=c("No", "Yes")) +
  labs(title = "RSE of Mean Life Satisfaction Score By Sexual Orientation",
       x = "Heterosexual or Not",
       y = "RSE of Mean Life Satisfaction Score") +
  theme_minimal()
```

# Result
The data of the overall life satisfaction rate analyzed different groups of people by sex, which contains 3 columns: Males, Females, Total persons respectively. Figure 1 shows the mean life satisfaction scores among different age groups: 15 – 24, 25 – 39, 40 – 54, 55 – 69, 70 years old and over respectively. It shows an obvious upward trend on the mean life satisfaction score with the age going higher (from a mean score of 6.9 of people aged from 15 – 34, to 7.9 of people aged 70 years or older. Figure two divided people into two groups by their migration status, which results in similar mean scores between Recent migrants and non-recent migrant. But non-recent migrants tend to have a higher mean score especially for females refer to the data. Figure 3 shows the mean scores of people with a mental health condition and people without a mental health condition (a score of 5.8 compared to 7.4). Same as mental health, people with a long-term physical health condition also tend to have a lower mean score of life satisfaction (Figure 4) Figure 5 shows that people with disability have a mean score of 6.7, compared to 7.4 of people without disability. Figure 6 shows the mean scores between homosexual and heterosexual people. We concluded that Heterosexual has a higher overall mean score than homosexual and bisexual. Within the male homosexual group, their life satisfaction score is as much as heterosexual. Figure 7 – 8 shows the relative standard error of mean, for the Figure 2 and Figure 6. Both shows that the RSE of two different groups differs to a large extent. A large RSE means that the scores widely spread around the population mean. For example, gay, lesbian or Bisexual has a much larger RSE than Heterosexual, so we can draw a conclusion that Heterosexual people have relatively concentrated life satisfaction scores.

Figure 9-10 showed the scores among several aggregated groups. By comparing the overall life satisfaction mean scores from different years (2014, 2019, 2020). We found that people’s overall scores decreased a bit in 2020 (From around 7.6 to 7.2). As of covid pandemic is introduced in australia in January 2020, people’s life satisfaction score fell by around 0.3 unit. 
\newpage

# Discussion

As the mean life satisfaction score goes lower while the age group goes younger, we observed a negative effect of aging. This is mainly because life satisfaction of older people is mainly affected by their life style. While younger people have a lower/decreasing life satisfaction, government should emphasize more on social/work benefits to reduce the living cost of younger people, as the outer happiness takes the mainly role in their life satisfaction. It is clear that the mean life satisfaction score is higher in the heterosexual group. But according to the RSE difference between the homosexual group and heterosexual group. We found that Australia has an unstable distribution of life satisfaction among different sexuality groups. It reflects that Australia needs to improve the life quality of homosexual group and try to eliminate the discrimination on the homosexual/bisexual group. People with physical or mental conditions or disabilities have a much lower mean life satisfaction score than normal people, and this gap is increasing by time. Government should pay more attention on deploying more social benefits to people with mental health problems and disables.

•	Weakness
Statistics weakness: …

This survey converted people’s life satisfaction into a number (life satisfaction score). This is apparently not enough for asssessing the overall satisfaction level. The satisfaction contains factors such as healthy physical being, inner happiness and social life.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References